---
title: "Case Study Codes"
author: "KICADDS23.1F-001"
date: "2024-03-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Uploading the Dataset
```{r}
data<-read.csv("C:/Users/msuma/OneDrive/Documents/NIBM/2nd Semester/Case Study/Health officers/data F.csv",header = TRUE)
```
```{r}
colnames(data)
```
Structure of the dataset
```{r}
str(data)
```
Renaming Columns
```{r}
#renaming columns
library(dplyr)
data<- data %>%
  rename('RDHS Division' = RDHS................Division,
         'AGSD Medical Officers' = Administrative.Grade..Senior.and.Deputy..Medical.Officers,
         'AGSD Non Medical Officers' = Administrative.Grade..Senior.and.Deputy..Non.Medical.Officers,
         'Specialists / Consultant' = Specialists.Consultant..other.than.administrative.grade.,
         'Hospital MO' = Hospital.Medical.Officers..D.M.O...M.O.I.C...S.H.O...S.M.O..in.OPD..etc..,
         'MO in MOH,AMOH' = Medical.Officers.in.MOH.AMOH,
         'School MO' = School.Medical.Officers,
         'Malaria MO' = Medical.Officers..Malaria.,
         'Leprosy MO' = Medical.Officers..Leprosy.,
         'Filaria MO' = Medical.Officers..Filaria.,
         'Principals/Sister Tutors ' = Principals.Sister.Tutors,
         'Venereal Diseases MO' = Medical.Officers..Venereal.Diseases.,
         'Tuberculosis MO' = Medical.Officers..Tuberculosis.,
         'Maternal and Child Health MO' =Medical.Officers..Maternal.and.Child.Health.,
         'Judicial MO' =Judicial.Medical.Officers,
         'Blood Bank MO' =Medical.Officers..Blood.Bank.,
         'Internee MO'=Internee.Medical.Officers..H.O..,
         'P.G.I.M.Trainees'=P.G.I.M..Trainees...,
         'Other MO'=Other.Medical.Officers,
         'Medical Officers' = Medical.Officers.1,
         'Total MO' = Total.Medical.Officers.2,
         'Consultant DS'=Consultant.Dental.Surgeons,
         'Regional DS'=Regional.Dental.Surgeons,
         'Dental Surgeons'=Dental.Surgeons,
         'Total DS' =Dental.Surgeons.3,
         'Ward Masters/Sisters'=Ward.Masters..Sisters,
         'PHNO'=Public.Health.Nursing.Officers,
         'Supervising and Public PHNS' =Supervising.Public.Health.Nursing.Sisters.Public.Health.Nursing.Sisters,
         'Pupil Nurses'=Pupil.Nurses,
         'Total Nurses'=Total.Nurses,
         'Supervising PHMW'=Supervising.Public.Health.Midwives,
         'PHMW'=Public.Health.Midwives,
         'Hospital Midwives'=Hospital.Midwives,
         'Total MidWives'=Total.Mid.Wives)
```


```{r}
data
```


```{r}
#checking for duplicated data 
duplicated(data)
```
```{r}
#uploading the population dataset
pop_data<-read.csv("C:/Users/msuma/OneDrive/Documents/NIBM/2nd Semester/Case Study/Health officers/Final population.csv",header = TRUE)
```

```{r}
#renaming a column
library(dplyr)

pop_data <- pop_data %>%
  rename(`RDHS Division` = RDHS................Division)
pop_data
```
Merging the two datasets
```{r}
#merging the two datasets
final_dataset<-merge(pop_data,data,by='RDHS Division')
final_dataset
```
```{r}
#examining the structure of the final dataset
str(final_dataset)
```
Population bar graph
```{r}

library(RColorBrewer)

# Create a gradient color palette with shades of blue
blue_palette <- colorRampPalette(colors = c("darkblue", "lightblue", "green"))

# Determine the number of unique RDHS Divisions
num_divisions <- length(unique(final_dataset$`RDHS Division`))

# Creating bar plot
barplot(final_dataset$population, 
        main = "Regional Population Distribution", 
        col = blue_palette(num_divisions),
        names.arg = final_dataset$`RDHS Division`,
        las = 2,
        cex.names = 0.76,
        ylim = c(0,2500000))
```
plot to visualize the distribution of medical officers by type
```{r}

library(ggplot2)
library(reshape2)


# Melt the data for plotting
melted_df <- melt(final_dataset[, c("RDHS Division", "Total MO", "Total DS", "Total Nurses")], id.vars = "RDHS Division")

# Plot stacked bar plot
ggplot(melted_df, aes(x = `RDHS Division`, y = value, fill = variable)) +
  geom_bar(stat = "identity") +
  labs(title = "Medical Officer Distribution by Type",
       x = "RDHS Division", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
Number of medical staff per 100,000 population
```{r}

#Number of Medical doctors
N_MO <- (sum(final_dataset$`Total MO`)/sum(final_dataset$population,na.rm = TRUE))*100000
N_MO
#Number of nurses
N_Nurses <-(sum(final_dataset$`Total Nurses`)/sum(final_dataset$population,na.rm = TRUE))*100000
N_Nurses
#Number of Dental surgeons
N_DS <-(sum(final_dataset$`Total DS`)/sum(final_dataset$population,na.rm = TRUE))*100000
N_DS
#Number of Public health midwives
N_MW <-(sum(final_dataset$PHMW)/sum(final_dataset$population,na.rm = TRUE))*100000
N_MW
```
```{r}
#summary of final dataset
summary(final_dataset)
```
Correlation heatmap
```{r}

population_total_all <- final_dataset[, c("population", "Total MO", "Total MidWives", "Total Nurses", "Total DS")]

# Calculating correlation matrix
correlation_matrix_all <- cor(population_total_all)

# Plotting the correlation heatmap
library(ggplot2)
library(reshape2)

# Melt correlation matrix for visualization
melted_corr_matrix_all <- melt(correlation_matrix_all)

# Create heatmap with correlation values
ggplot(data = melted_corr_matrix_all, aes(x = Var1, y = Var2, fill = value, label = round(value, 2))) +
  geom_tile() +
  geom_text(color = "black") +  # Add text labels
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1,1), name = "Correlation") +
  labs(title = "Correlation Heatmap between Population and Health Staff", x = "Variables", y = "Variables")

```
```{r}
#Number of Medical Officers
library(readxl)
RCBD_MO <- read_excel("C:/Users/msuma/OneDrive/Documents/NIBM/2nd Semester/Case Study/Health officers/RCBD datasets/RCBD MO.xlsx",col_types = c("text", "text", "numeric"))
```

Normality tests on medical officers
```{r}

shapiro.test(final_dataset$`Specialists / Consultant`)
shapiro.test(final_dataset$`Hospital MO`)
shapiro.test(final_dataset$`MO in MOH,AMOH`)
shapiro.test(final_dataset$`School MO`)
shapiro.test(final_dataset$`Malaria MO`)
shapiro.test(final_dataset$`Filaria MO`)
shapiro.test(final_dataset$`Leprosy MO`)
shapiro.test(final_dataset$`Venereal Diseases MO`)
shapiro.test(final_dataset$`Tuberculosis MO`)
shapiro.test(final_dataset$Epidemiologists)
shapiro.test(final_dataset$`Maternal and Child Health MO`)
shapiro.test(final_dataset$`Judicial MO`)
shapiro.test(final_dataset$`Blood Bank MO`)
shapiro.test(final_dataset$`Internee MO`)
shapiro.test(final_dataset$P.G.I.M.Trainees)
shapiro.test(final_dataset$`Other MO`)
```

Fried Man test on medical officers
```{r}
library(agricolae)
G<-friedman(RCBD_MO$Position,RCBD_MO$Region,RCBD_MO$Count)
G
```

visualization of groups according to the distribution of medical officers
```{r}
plot(G)
```
```{r}
#Number of Nurses
library(readxl)
RCBD_NS <- read_excel("C:/Users/msuma/OneDrive/Documents/NIBM/2nd Semester/Case Study/Health officers/RCBD datasets/RCBD NS.xlsx")
```

Normality test on nurses
```{r}
shapiro.test(final_dataset$Matrons)
shapiro.test(final_dataset$`Ward Masters/Sisters`)
shapiro.test(final_dataset$`Principals/Sister Tutors `)
shapiro.test(final_dataset$Nursing.Officers)
shapiro.test(final_dataset$PHNO)
shapiro.test(final_dataset$`Supervising PHMW`)
shapiro.test(final_dataset$`Pupil Nurses`)
```

Fried Man test on Nurses
```{r}
library(agricolae)
H<-friedman(RCBD_NS$Position,RCBD_NS$Region,RCBD_NS$Count)
H
```
visualization of groups according to the distribution of nurses
```{r}
plot(H)
```
```{r}
library(readxl)
RCBD_Dental <- read_excel("C:/Users/msuma/OneDrive/Documents/NIBM/2nd Semester/Case Study/Health officers/RCBD datasets/RCBD Dental.xlsx")
```

Normality test on dental surgeons
```{r}
shapiro.test(final_dataset$`Consultant DS`)
shapiro.test(final_dataset$`Regional DS`)
shapiro.test(final_dataset$`Dental Surgeons`)
shapiro.test(final_dataset$P.G.I.M.Trainees..)
```

Fried Man test on Dental Surgeons
```{r}
library(agricolae)
J<-friedman(RCBD_Dental$Position,RCBD_Dental$Region,RCBD_Dental$Count)
J
```
Visualization of groups according to the distribution of dental surgeons
```{r}
plot(J)
```

